---
title: "Lab 05"
author: "Elyse McCormick"
date: "`r Sys.Date()`"
output: html_document
subtitle: Analysis of Env. Data
---

**Q1 (2 pts.): Show the R code you used to create exp_fun()**

```{r}
exp_fun = function(x, a, b) 
{
  return(a * exp(-b * x))
}
curve(
  exp_fun(x, 1, -1/10), from = 0, to = 10, add = FALSE,
  main = "Exponential Function Test Plot",
  xlab = "x", ylab = "f(x)")
```

**Q2 (4 pts.): In your lab report, include a single figure containing four negative exponential curves with different parameter values and line colors/textures.**

```{r}
curve(
  exp_fun(x, 1.9, 0.1), col = "black",  
  from = 0, to = 50, add = FALSE, main = "Exponential Function Test Plot",
  xlab = "x", ylab = "f(x)")
curve(
  exp_fun(x, 1.9, 0.3), col = "black", lty = "dotted", 
  from = 0, to = 50, add = TRUE)
curve(
  exp_fun(x, 1.2, 0.2), col = "red", 
  from = 0, to = 50, add = TRUE)
curve(
  exp_fun(x, 1.2, 0.4), col = "red", lty = "dotted", 
  from = 0, to = 50, add = TRUE)
```

**Q3 (2 pts.): Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter a**

As you vary parameter a, you see a different starting point on the y axis. For example, when a = 1.9, the black lines start near the top of the y axis. But when a = 1.2, the line starts near the midpoint of the y axis. 

**Q4 (2 pts.): Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter b**

As you vary parameter b, it shifts the rate of decay, or the slope of the curve. For the black lines, when b = 0.1 the curve is more shallow than when b = 0.3. Similarly, for the red lines, when b = 0.2 the curve is more shallow than when b = 0.4

**Q5 (6 pts.): In your lab report, include a single plot containing 6 Ricker curves with different parameter values.**

```{r}
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}
curve(
  ricker_fun(x, 25, 0.2), col = "black",
  from = 0, to = 50, ylim = c(0, 100), add = FALSE, main = "Ricker Functions Test Plot",
  ylab = "f(x)", xlab = "x")
curve(
  ricker_fun(x, 20, 0.2), col = "black", lty = "dotted",
  from = 0, to = 50, ylim = c(0, 100), add = TRUE, main = "Ricker Functions Test Plot",
  ylab = "f(x)", xlab = "x")
curve(
  ricker_fun(x, 10, 0.2), col = "black", lty = "dotted",
  from = 0, to = 50, ylim = c(0, 100), add = TRUE, main = "Ricker Functions Test Plot",
  ylab = "f(x)", xlab = "x")
curve(
  ricker_fun(x, 75, 0.3), col = "red", 
  from = 0, to = 50, ylim = c(0, 100), add = TRUE, main = "Ricker Functions Test Plot",
  ylab = "f(x)", xlab = "x")
curve(
  ricker_fun(x, 50, 0.3), col = "red", lty = "dotted",
  from = 0, to = 50, ylim = c(0, 100), add = TRUE, main = "Ricker Functions Test Plot",
  ylab = "f(x)", xlab = "x")
curve(
  ricker_fun(x, 40, 0.3), col = "red", lty = "dotted",
  from = 0, to = 50, ylim = c(0, 100), add = TRUE, main = "Ricker Functions Test Plot",
  ylab = "f(x)", xlab = "x")
```
**Q6 (2 pts.): Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter a**

As you vary parameter a, the slope of the curve varies. The highest peak exists for when a = 75, the lowest peak of the curves exists when a = 10. 

**Q7 (2 pts.): Observe how the curves vary as you change the two parameters’ values. Qualitatively describe what happens to the curve as you vary parameter b**

As you vary parameter b, the y intercept of the curve varies. As b decreases, the change in the curve shrinks, and works with a to flatten the shape of the curve. 

**Q8 (2 pts.): Linear Model. Provide the values of the slope, x1, and y1 parameters you chose. Briefly describe how you chose the values.**

x1 = 1500
y1 = 0
slope = -0.00025

The first thing I did was just visually assess what might be a decent fit for the data, approximating that a linear model might be "pulled up" from a flat slope by the points around y = 0.6 and 0.8, and then trying to visually see where the line would run in the middle of the most points. I chose these values because I wanted the line to cross the y axis at 0, and have a slope that ran through the middle of the point range. I toyed with the slope until it looked similar to what I wanted. 

**Q9 (2 pts.): In your lab report, include a scatterplot of the salamander data with your fitted linear model.**
```{r echo=TRUE}
require(psych)
require(here)
sal_dat = read.csv(here("data", "dispersal.csv"))

plot(
  sal_dat$dist.class,
  sal_dat$disp.rate.ftb,
  xlim = c(0, 1500),
  xlab = "Distance Class", 
  ylab = "Standardized Dispersal Rate", 
  main = "Marbled Salamander - First Time Breeder Approximate Fit")
line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}
curve(line_point_slope(x, 1500, 0, -0.00025), add = TRUE)
```

**Q10 (2 pts.): Exponential Model. Provide the values of the a and b. Briefly describe how you chose the values.**

a = 0.7
b = 1/300

I chose these values purely by trial and error of what might make the curve fit the data points best. Since a is where the curve crosses the y axis, you can play with b to see how to affect the shape of the curve. I played with the a and b values until they worked together to make a line that approximated the middle of the data. 

**Q11 (2 pts.): In your lab report, include a scatterplot of the salamander data with your fitted exponential model.**

```{r echo=TRUE}
plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander - First Time Breeders Exponential Model")
exp_fun = function(x, a, b) 
     {
       return(a * exp(-b * x))
     }
curve(exp_fun(x, 0.7, 1/300), from = 0, to = 1500, add = TRUE)
```

**Q12 (2 pts.): Ricker Model Provide the values of the a and b. Briefly describe how you chose the values.**

a = 0.03
b = 1/300
I chose a and b parameters through trial and error of what made normal looking curve that fit the graph. Since 1/b is where the peak of the curve is on the x-axis, you can move the parameter of a to adjust the initial slope to fit the data. This seems like a very bad fit for the data.

**Q13 (2 pts.): In your lab report, include a scatterplot of the salamander data with your fitted Ricker model.**

```{r echo=TRUE}
plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander - First Time Breeders Ricker Model")
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}
curve(
  ricker_fun(x, 0.003, 1/300), from = 0, to = 1500, add = TRUE)
```

**Q14 (4 pts.): Show the R code you used to create your data frame of model residuals.**
```{r}
guess_x = 1500
guess_y = 0
guess_slope = -0.00025

plot(
  sal_dat$dist.class,
  sal_dat$disp.rate.ftb,
  xlim = c(0, 1500),
  xlab = "Distance Class", 
  ylab = "Standardized Dispersal Rate", 
  main = "Marbled Salamander - first time breeders\n(Bad) linear model")
line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}
curve(line_point_slope(x, 1500, 0, -0.00025), add = TRUE)

line_point_slope(sal_dat$dist.class, guess_x, guess_y, guess_slope)

sal_dat$y_predicted = line_point_slope(sal_dat$dist.class, guess_x, guess_y, guess_slope)

sal_dat$resids_linear = sal_dat$disp.rate.ftb - sal_dat$y_predicted

sum(sal_dat$resids_linear, na.rm = TRUE)

#plot( x = sal_dat$disp.rate.ftb, xlab = "Observed", y = sal_dat$resids, ylab = "Residuals", pch = 8)

#hist(sal_dat$resids_linear, xlab = "Residuals", main = "Histogram of Linear Residuals")

guess_xe = 300
guess_a = 0.7
guess_b = 1/300

plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander - First Time Breeders Exponential Model")
exp_fun = function(x, a, b) 
{
  return(a * exp(-b * x))
}
curve(exp_fun(x, 0.7, 1/300), from = 0, to = 1500, add = TRUE)

line_point_slope(sal_dat$dist.class, guess_x, guess_a, guess_b)

sal_dat$ey_predicted = line_point_slope(sal_dat$dist.class,guess_xe, guess_a, guess_b)

sal_dat$resids_exp = sal_dat$disp.rate.ftb - sal_dat$ey_predicted

sum(sal_dat$resids_exp, na.rm = TRUE)

sum(abs(sal_dat$resids_exp))

#plot( x = sal_dat$disp.rate.ftb, xlab = "Observed", y = sal_dat$resids_exp, ylab = "Residuals", pch = 8)

#hist(sal_dat$resids_exp, xlab = "Residuals", main = "Histogram of Exp Residuals")

guess_xr = 300
guess_ar = 0.003
guess_br = 1/300

plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander - First Time Breeders Ricker Model")
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}
curve(
  ricker_fun(x, 0.003, 1/300), from = 0, to = 1500, add = TRUE)
line_point_slope(sal_dat$dist.class, guess_xr, guess_ar, guess_br)

sal_dat$ry_predicted = line_point_slope(sal_dat$dist.class,guess_xr, guess_ar, guess_br)

sal_dat$resids_ricker = sal_dat$disp.rate.ftb - sal_dat$ry_predicted

sum(sal_dat$resids_ricker, na.rm = TRUE)

sum(abs(sal_dat$resids_ricker))

#plot( x = sal_dat$disp.rate.ftb, xlab = "Observed", y = sal_dat$resids_ricker, ylab = "Residuals", pch = 8)

par(mfrow = c(1,3))
plot(
  sal_dat$dist.class,
  sal_dat$disp.rate.ftb,
  xlim = c(0, 1500),
  xlab = "Distance Class", 
  ylab = "Standardized Dispersal Rate", 
  main = "Marbled Salamander - first time breeders\n(Bad) linear model")
line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}
curve(line_point_slope(x, 1500, 0, -0.00025), add = TRUE)
plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander - First Time Breeders Exponential Model")
exp_fun = function(x, a, b) 
{
  return(a * exp(-b * x))
}
curve(exp_fun(x, 0.7, 1/300), from = 0, to = 1500, add = TRUE)
plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander - First Time Breeders Ricker Model")
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}
curve(
  ricker_fun(x, 0.003, 1/300), from = 0, to = 1500, add = TRUE)


```


**Q15 (3 pts.): In your lab report, include histograms of the residuals for each of your three models. You may create a single figure with three panels, or include three separate figures.**

```{r}
par(mfrow = c(1, 3))
hist(sal_dat$resids_linear, xlab = "Residuals", main = "Histogram of Linear Residuals", col = "red")
hist(sal_dat$resids_exp, xlab = "Residuals", main = "Histogram of Exp Residuals", col = "yellow")
hist(sal_dat$resids_exp, xlab = "Residuals", main = "Histogram of Ricker Residuals", col = "blue")
```




