---
title: 'Lab 12: Beyond the General Linear Model'
author: "Elyse McCormick"
date: "`r Sys.Date()`"
output: html_document
subtitle: "Analysis of Environmental Data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Q1 (2 pts.): Include your plot in your lab report.**

```{r}

require(here)

# Read in data

dat_bird = read.csv(
  here("data", "bird.sub.csv")
)
head(dat_bird)

dat_hab = read.csv(
  here("data", "hab.sub.csv")
)
head(dat_hab)

birdhab <- data.frame(merge(dat_bird, dat_hab))
head(birdhab)


alpha = 0.05
n_sims = 1000
int_obs = 0.0991
slope_obs = 0.00584
sd_obs = 0.1412668

linear = function(x, y_int, slope)
{
      return(y_int + x * slope)
}

linear_simulator = function(x, y_int, slope, st_dev)
{
  y <- linear(x, y_int, slope)
  n <- length(x)
  noise <- rnorm(n, mean = 0, sd = st_dev)
  
    return(y + noise)
}

linear_sim_fit = function(x, slope, y_int, st_dev)
{
  y_sim = linear_simulator(
    x = x,
    y_int = y_int,
    slope = slope,
    st_dev = st_dev
  )
  fit_sim = lm(y_sim ~ x)
  return(fit_sim)
}

p_vals = numeric(n_sims)

# specify the number of different standard deviation values to simulate:
n_sds = 20
pop_sds = seq(from = 0.01, to = 1.5, length.out = n_sds)

pop_sd_powers = numeric(n_sds)

for(j in 1:n_sds)
{
    for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(x = birdhab$ls,
                             y_int = int_obs,
                             slope = slope_obs,
                             st_dev = pop_sds[j])
    p_vals[i] = summary(fit_sim)$coefficients[2, "Pr(>|t|)"]
  }
  pop_sd_powers[j] = sum(p_vals < alpha) / n_sims
}

sim_output_dispersion = data.frame(
  sd = pop_sds,
  power = pop_sd_powers)

fit_lowess_30 = loess(power ~ sd, data = sim_output_dispersion, span = 0.3)
newdata_disp = data.frame(sd = seq(0, 1.5, length.out = 100)) 
plot(
  x = newdata_disp$sd,
  y = predict(fit_lowess_30, newdata = newdata_disp),
  type = "l",
  ylab = "Statistical Power", xlab = "Dispersion",
  main = "Dispersion vs. Power LOWESS/original"
  )
points(x = sim_output_dispersion$sd, y = sim_output_dispersion$power, col = "red")

legend("bottomleft",
       c("original \n","smooth"), 
       lty=c(NA,1),pch=c('o',NA),
       horiz=TRUE, cex=0.6)
```


**Q2 (2 pts.): Include your plot in your lab report.**

```{r}
require(psych)
require(here)
sal_dat = read.csv(here("data", "dispersal.csv"))

plot(sal_dat$dist.class,
     sal_dat$disp.rate.ftb,
     xlim = c(0, 1500),
     xlab = "Distance Class", 
     ylab = "Standardized Dispersal Rate", 
     main = "Marbled Salamander Dispersal")
ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}
curve(
  ricker_fun(x, 0.008, 0.004), from = 0, to = 1500, add = TRUE)

fit_ricker_nls = nls(
  disp.rate.ftb ~ ricker_fun(dist.class, a, b),
  data = sal_dat,
  start = list(b = 0, a = 1))
summary(fit_ricker_nls)

dist_newdata = data.frame(dist.class = seq(0, 1600, length.out = 1600))
lines(predict(fit_ricker_nls, newdata = dist_newdata, ), col = "blue")
legend("topright", 
       legend = c("nls", "manual"), 
       lty = 1, col = c("blue","black"))
```

**Q3 (1 pt.): What are the AIC values for each of the 4 models?**

```{r}
#Packages

require(here)

#Import Data

dat_bird = read.csv(here("data", "bird.sta.csv"))
dat_habitat = read.csv(here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

dat_all$GCKI_pres = dat_all$GCKI > 0

# Create model fits
fit_GCKI_slope = glm(GCKI_pres ~ slope, data = dat_all, family = binomial)
fit_GCKI_ba_tot = glm(GCKI_pres ~ ba.tot, data = dat_all, family = binomial)
fit_GCKI_both_additive = glm(GCKI_pres ~ slope + ba.tot, data = dat_all, family = binomial)
fit_GCKI_both_interactive = glm(GCKI_pres ~ slope * ba.tot, data = dat_all, family = binomial)

n = 500

slope_newdata = data.frame(
  slope = seq(
    from = min(dat_all$slope, na.rm = T),
    to = max(dat_all$slope, na.rm = T),
    length.out = n
  )
)

ba_newdata = data.frame(
  ba.tot = seq(
    from = min(dat_all$ba.tot, na.rm = T),
    to = max(dat_all$ba.tot, na.rm = T),
    length.out = n
  )
)

slope_newdata$GCKI_predicted = 
  predict(
    fit_GCKI_slope,
    newdata = slope_newdata,
    type = "response"
  )

ba_newdata$GCKI_predicted = 
  predict(
    fit_GCKI_ba_tot,
    newdata = ba_newdata,
    type = "response"
  )

# Plot parameters

AIC(
  fit_GCKI_ba_tot,
  fit_GCKI_slope,
  fit_GCKI_both_additive,
  fit_GCKI_both_interactive)

```

**Q4 (1 pt.): Which model would you choose, and why?**

I would choose the model for the Golden Crowned Kinglet with both total basal area and slope interacting (fit_GCKI_both_interactive). Based on it's low AIC score, that indicates that it is the best fit for the model, and thus the model that best explains the data. 

**Q5 (1 pt.): Based on the model coefficient table of your chosen model, describe the direction and significance of the relationship(s) of the predictor variable or variables to the binary response. Make sure your answer is in terms of the ecological context.**

```{r}

fit_GCKI_both_interactive = glm(GCKI_pres ~ slope * ba.tot, data = dat_all, family = binomial)
summary(fit_GCKI_both_interactive)
```
There is a significant effect of both total basal area and the interaction between slope and basal area. This means that together, slope and basal area negatively impact Golden Crowned Kinglet presence in this area. 

**Q6 (2 pts.): Include your two single-predictor model plots in your report.**

```{r}
par(mfrow = c(2, 1))

# Presence/absence data, translucent points:
plot(
  GCKI_pres ~ slope, data = dat_all,
  xlab = "Percent Slope",
  ylab = "GCKI presence/absence",
  pch = 16, cex = 1.5, col = gray(0, 0.2)
)

lines(GCKI_predicted ~ slope, data = slope_newdata)

plot(
  GCKI_pres ~ ba.tot, data = dat_all,
  xlab = "Basal Area",
  ylab = "GCKI presence/absence",
  pch = 16, cex = 1.5, col = gray(0, 0.2)
)

lines(GCKI_predicted ~ ba.tot, data = ba_newdata)
```


**Q7 (4 pts.): Include contour plots (or interactive 3D perspective plots) in your report.**

```{r}

require(rgl)

n = 50

ba.tot = seq(
  from = min(dat_all$ba.tot, na.rm = T),
  to = max(dat_all$ba.tot, na.rm = T),
  length.out = n)
slope = seq(
  from = min(dat_all$slope, na.rm = T),
  to = max(dat_all$slope, na.rm = T),
  length.out = n)

# I can take advantage of the expand.grid() function to make a data.frame 
# that contains every combination of the two predictors:

new_dat_all = expand.grid(
  ba.tot = ba.tot,
  slope = slope)
head(new_dat_all)

tail(new_dat_all)

new_dat_all$pred_add = predict(
  fit_GCKI_both_additive,
  newdata = new_dat_all,
  type = "response")

z_GCKI_add = matrix(
  new_dat_all$pred_add,
  nrow = length(ba.tot),
  byrow = FALSE)

# Create a color ramp function using heat colors
colorfunc = colorRampPalette(
  heat.colors(length(ba.tot)))

# Figure out the indices of the color ramp
col_indices_GCKI_add = findInterval(
  new_dat_all$pred_add,
  seq(
    min(new_dat_all$pred_add),
    max(new_dat_all$pred_add),
    length.out = 50))

rgl::persp3d(
  x = ba.tot,
  y = slope,
  z = z_GCKI_add,
  xlab = "Basal Area",
  ylab = "Slope",
  zlab = "Pr(present)",
  alpha = 0.9,
  col = colorfunc(50)[col_indices_GCKI_add]
)
rglwidget()
```


