---
title: "Final R Reference Guide"
author: "Elyse McCormick"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}

## Loading Data and Packages

**Loading data and packages**

**Loading packages**

Both require() and library() will load packages

```{r}
require(here)
library(palmerpenguins)
```

**Loading Data**

```{r}
ginkgo_dat = read.csv(
  here("data", "ginkgo_data_2022.csv")
)
head(ginkgo_dat)
```

## Data structures

**Data Structures**

**c() function**

The function c() combines or concatenates its arguments into a vector(a 1-dimensional data structure consisting of 1 or more elements). All of the elements must be of the same type, i.e. you can’t combine character and numeric types in the same call to c(). 

Here’s examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

```

I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)
```
**length() function**

The length() function can be used to get or set the length of vectors, lists, and factors, or any other R object that you've defined. 

```{r}
length(num_vec)
```
**matrix() function**

The matrix() function creates a matrix from a given set of data or values. 

```{r}
# nrow is the number of rows you want
# ncol is the number of columsn you want
# byrow is a logical argument. If it's FALSE, then your matrix gets filled by the column. TRUE populates by row.

matrix(data = num_vec, nrow = 2, ncol = 2, byrow = FALSE)
```
**data.frame()**

The data.frame() function creates a data frame of your data, which is tightly coupled collections of data in rows and columns that are used as the base data structure in R. 

```{r}
ginkgo_data <- data.frame(ginkgo_dat)
```

**nrow() and ncol() functions**

Both nrow() and ncol() give you the number of rows or columns present in x. 

```{r}
nrow(ginkgo_data)
ncol(ginkgo_data)
```

**dim() function**

The dim() function gives the dimensions of an object or data frame.

```{r}
dim(ginkgo_data)
```
## Subsetting

**Subsetting**

You can subset a dataframe by using the $ to pull out a specific column of data. 

```{r}
ginkgo_data$max_width
```
You can also subset by position using square brackets. 

```{r}
# select first row of the ginkgo data
ginkgo_data[1,]

# select the element in row 2, column 3
ginkgo_data[2,3]

#select the 3rd column of the ginkgo data
ginkgo_data[,3]
```

Finally, you can use the subset function to pull specific aspects of your data. For example, if I wanted only the information about Adelie penguins from the palmerpenguins dataset, I could do that using the subset function!

```{r}
require(palmerpenguins)
dat_pen = droplevels(subset(penguins, species == "Adelie"))
```

## Numerical Data Exploration

**Numerical Data Exploration**

**summary() function**

The summary() function will give you a basic overview of your selected object. In a data frame, it will give you the overview summary statistics for each column.  

```{r}
summary(ginkgo_data)
```
**mean() function**

The mean() function will calculate the arithmetic mean of any group of numbers. 

```{r}
mean(ginkgo_data$max_depth)
```
**sd() function**

The sd() function will calculate the standard deviation of any group of numbers. 

```{r}
sd(ginkgo_data$max_depth)
```

## Graphical Data Exploration 

**Graphical Data Exploration**

**Creating a Scatterplot**

To create a scatterplot, you need to use the plot() function. Within this, you can specify what you want the x and y axis to be by subsetting the data and calling them "x" and "y". You can use  'xlim' and 'ylim' callss to specify a cutoff value for the x and y axes, here I called it 'NULL' since I didn't want to truncate any values. The calls 'xlab' and 'ylab' allow you to name the axes. The argument 'main' allows you to name the title of the plot. 'pch' allows you to choose an icon to plot as a data point, 'cex' allows you to set the size of the font used, and 'col' allows you to change the color of the data points. 
```{r}
plot(x = ginkgo_data$max_depth, y = ginkgo_data$max_width,
     xlim = NULL, ylim = NULL, xlab = "Maximum Leaf Depth (cm)", 
     ylab = "Maximum Leaf Width (cm)", main = "Ginkgo Leaf Dimensions",
     pch = 8, cex = 1, col = "gold")
```


**Creating a Histogram**

The hist() function allows you to create a histogram. First, you can specify which portion of your dataset you'd like to examine by subsetting a column. Then, you can specify the number of bins using the 'breaks' call. Then, you can specify the other aesthetic portions of the plot using the same methods as above. 

```{r}
require(palmerpenguins)
dat_pen <- palmerpenguins::penguins
hist(dat_pen$flipper_length_mm, breaks = 5, xlab = "Flipper Length (mm)", main = "Histogram of Penguin Flipper Length ", 
     col = "skyblue")
```


**Creating a Boxplot**

The boxplot() function allows you to create a boxplot. First, you can specify which portion of your dataset you'd like to examine by subsetting a column. Then, you can specify the other aesthetic portions of the plot using the same methods as above. To make a conditional boxplot, simply note the dependent variable as explained by the independent variable by using a tilde (~)

```{r}
par(mfrow = c(1,2))
boxplot(ginkgo_data$petiole_length, main = "Ginkgo Petiole Lengths", xlab = "", ylab = "Petiole Length (cm) ", col = "springgreen3")

boxplot(formula = max_width ~ seeds_present, data = ginkgo_data, main = "Leaf Width by Seed Presence",
        xlab = "Seed Presence", ylab = "Maximum Width (cm)", col = "gold")
```

**Creating a 2x2 Histogram Plot**

To make a multipanel plot, you can use the par(mfrow = ) argument to make a set of plot pairs. By listing what you need within the c function, you can make a 2 x 2 plot. 

```{r}
par(mfrow = c(2,2))

hist(dat_pen$body_mass_g, breaks = 5, xlab = "Body Mass (g)", main = "Histogram of Penguin Body Mass", 
     col = "midnightblue")

hist(dat_pen$bill_length_mm, breaks = 5, xlab = "Bill Length (mm)", main = "Histogram of Penguin Bill length ", 
     col = "skyblue")

hist(dat_pen$bill_depth_mm, breaks = 5, xlab = "Bill Depth (mm)", main = "Histogram of Penguin Bill Depth ", 
     col = "royalblue1")

hist(dat_pen$flipper_length_mm, breaks = 5, xlab = "Flipper Length (mm)", main = "Histogram of Penguin Flipper length ", 
     col = "steelblue")
```

## Distribution Functions

**Distribution Functions**

There are four distribution functions for normal distributions. They are as follows:

**dnorm():** the probability density, where x is the sample size, mean is the population average, and the sd is the standard deviation.

```{r}
dnorm(x = 2, mean = 2, sd = 4)
```

**pnorm():** calculates the cumulative probability density function. This is where you can calculate the area to the left of a curve, unless you go in and tell it to do something different. Here again, x is the sample size, mean is the population average, and the sd is the standard deviation. You can set the lower tail to either be true (the value less than the quartile) or false (the value greater than the quartile), but it is FALSE by default. 


```{r}
pnorm(q = 1.5, mean = 0, sd = 2, lower.tail = FALSE)
```

**qnorm():** this is the the quantile function, which calculates the inverse of the previous function. Here, p is the inverse, which is your value of interest, the x is the sample size, mean is the population average, and the sd is the standard deviation. Just as before, you can set the lower tail to either be true (the value less than the quartile) or false (the value greater than the quartile), but it is FALSE by default. 

```{r}
qnorm(p = 0.5, mean = 1, sd = 2, lower.tail = FALSE)
```


There are also three distribution functions for binomial distributions.

dbinom() gives you the probability density function of a normal binomial distribution, where x is the total number of successes, size is the total number of trials, and prob is the probability of success. 

```{r}
dbinom(x = 7, size = 8, prob = .5)
```

pbinom(): This gives you the area under the left side of the tail of a binomial distribution. This also helps you calculate the probability of an event. Here, x is the total number of successes, size is the total number of trials, and prob is the probability of success. You can set the lower tail to either be true (the value less than the quartile) or false (the value greater than the quartile), but it is FALSE by default. 

This give you the value to the left of your p value. 

```{r}
#pbinom() used to find 7 or fewer successes, when the sample size is 30 items, and the probability is 30%
pbinom(7, size=30, prob=.3, lower.tail = FALSE)
```

qbinom(): This gives the area under the binomial distribution. This helps calculate the probability of an event to the 'p'th quantile. Here, x is the total number of successes, size is the total number of trials, and prob is the probability of success. You can set the lower tail to either be true (the value less than the quartile) or false (the value greater than the quartile), but it is FALSE by default. 

This gives you the value to the left of your q value. 

```{r}
#qbinom() used to find the 25th quantile in 30 trials with a probability of 40%
qbinom(.25, size=30, prob=.4)

```





















